---
title: "R Notebook"
output: html_notebook
---



```{r}
#packages needed
library(ggplot2)
library(grid)
library(gridExtra)
library(gtable)
```

# 1 combine two graphs with only one x-axis

We first create 3 figures using the dataset 'OrchardSprays)
```{r}
# figure without x-axis and labels
fig1 <- ggplot(OrchardSprays, aes(treatment, decrease, colour=factor(treatment)))+
  geom_boxplot()+
  ylim(ymin=0, ymax=150)+
  theme(legend.position='none',
        axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())



fig2 <- ggplot(OrchardSprays, aes(treatment, rowpos*colpos, colour=factor(treatment)))+
  geom_boxplot()+
  ylim(ymin=0, ymax=150)+
  theme(legend.position='none')

fig3 <- ggplot(OrchardSprays, aes(treatment, colpos, colour=factor(treatment)))+
  geom_boxplot()
```


# arrange figures via gtable

The problem in the resulting graph is the wrong scaling of the graph area
```{r}
g1 <- ggplotGrob(fig1) 
g2 <- ggplotGrob(fig2)

mat <- matrix(list(g1,g2), nrow=2)
a <- gtable_matrix(name = "demo", grobs = mat, 
                   widths = unit(c(12), "cm"), 
                   heights = unit(c(3, 3), "cm"))
grid.newpage()
grid.draw(a)
```


The previous combined graph does not look nice because the chart area is not the same.
To solve this I found a chunk of code on the internet
Just run this code chunk to load the functions

```{r}
# define panel.hist function based on the code from ?pairs

panel.hist <- function(x, ...)
{
	usr <- par("usr"); on.exit(par(usr))
	par(usr = c(usr[1:2], 0, 1.5) )
	h <- hist(x, plot = FALSE)
	breaks <- h$breaks; nB <- length(breaks)
	y <- h$counts; y <- y/max(y)
	rect(breaks[-nB], 0, breaks[-1], y, col = "cyan", ...)
}

# define panel.cor function based on the code from ?pairs

panel.cor <- function(x, y, digits = 2, prefix = "", cex.cor, ...)
{
	usr <- par("usr"); on.exit(par(usr))
	par(usr = c(0, 1, 0, 1))
	r <- abs(cor(x, y))
	txt <- format(c(r, 0.123456789), digits = digits)[1]
	txt <- paste0(prefix, txt)
	if(missing(cex.cor)) cex.cor <- 0.8/strwidth(txt)
	text(0.5, 0.5, txt, cex = cex.cor * r)
}

# function to arrange plots borrowed from "https://github.com/baptiste/gridextra/wiki/arranging-ggplot#fix-the-panel-size"

gtable_frame <- function(g, width=unit(1,"null"), height=unit(1,"null")){
  panels <- g[["layout"]][grepl("panel", g[["layout"]][["name"]]), ]
  ll <- unique(panels$l)
  tt <- unique(panels$t)
  
  fixed_ar <- g$respect
  if(fixed_ar) { # there lies madness, we want to align with aspect ratio constraints
    ar <- as.numeric(g$heights[tt[1]]) / as.numeric(g$widths[ll[1]])
    print(ar)
    height <- width * ar
    g$respect <- FALSE
  }
  
  core <- g[seq(min(tt), max(tt)), seq(min(ll), max(ll))]
  top <- g[seq(1, min(tt)-1), ]
  bottom <- g[seq(max(tt)+1, nrow(g)), ]
  left <- g[, seq(1, min(ll)-1)]
  right <- g[, seq(max(ll)+1, ncol(g))]
  
  fg <- nullGrob()
  lg <-  if(length(left))  g[seq(min(tt), max(tt)), seq(1, min(ll)-1)] else fg
  rg <- if(length(right)) g[seq(min(tt), max(tt)), seq(max(ll)+1,ncol(g))] else fg
  grobs = list(fg, g[seq(1, min(tt)-1), seq(min(ll), max(ll))], fg, 
               lg, g[seq(min(tt), max(tt)), seq(min(ll), max(ll))], rg, 
               fg, g[seq(max(tt)+1, nrow(g)), seq(min(ll), max(ll))], fg)
  widths <- unit.c(sum(left$widths), width, sum(right$widths))
  heights <- unit.c(sum(top$heights), height, sum(bottom$heights))
  all <- gtable_matrix("all", grobs = matrix(grobs, ncol=3, nrow=3, byrow = TRUE), 
                       widths = widths, heights = heights)
  all[["layout"]][5,"name"] <- "panel" # make sure knows where the panel is
  if(fixed_ar)  all$respect <- TRUE
  all
}

```


Now plot the two graphs again but first rescale them using the functions defined in the chunk above.
```{r}
g1 <- ggplotGrob(fig1) 
g2 <- ggplotGrob(fig2)

fg1 <- gtable_frame(g1) #using function defined in the previous chunk
fg2 <- gtable_frame(g2) #using function defined in the previous chunk
fg12 <- rbind(fg1,fg2, size = "last")
grid.newpage()
grid.draw(fg12)

#setwd("D:/Bksercu/OneDrive - UGent/WE11c112/Bksercu/PhD/data/analysis/light_availability_paper")
#ggsave(file="plots/tree_shrub_cover_review_numbered.tiff", width = 20, height = 20, units = "cm",dpi=300, fg12)
```


# 2. put group factor as label on the axis
It might happen that you have two grouping factors and you want to plot boxplots per group for each group 2, ending up with indication of two grouplabels on your axis.
This can be done with facet_wrap using labels (I also showed it with flipped coordinates)
```{r}

labels <- c("1" = "Block 1", "2"= "Block 2", "3" = "Block 3", "4"= "Block 4", "5"= "Block 5", "6"= "Block 6")

ggplot(npk, aes(N, yield, colour=factor(N)))+
  geom_boxplot()+
  labs(title="block effects on yield for N-treatment", y= "yield")+
  facet_wrap(~block,ncol=6,strip.position = "bottom", scales = "free_x",
             labeller=labeller(block = labels))+
  theme(axis.text=element_text(size=14, face="bold", angle=0),
        axis.title=element_text(size=16,face="bold"),
        axis.title.x=element_blank(),
        axis.title.y=element_blank(),
        title=element_text(size=18,face="bold"),
        plot.title = element_text(hjust = 0.5),
        panel.spacing = unit(0, "lines"), 
        strip.background = element_blank(),
        strip.placement = "outside", 
        strip.text =element_text(size=14, face="bold", angle=0))

# horizontal boxplots with coord_flip()
fig4 <- ggplot(npk, aes(N, yield, colour=factor(N)))+
  geom_boxplot()+
  coord_flip()+
  labs(title="block effects on yield for N-treatment", y= "yield")+
  facet_wrap(~block,nrow=3,strip.position = "left", scales = "free_y",
             labeller=labeller(block = labels))+
  theme(axis.text=element_text(size=14, face="bold", angle=0),
        axis.title=element_text(size=16,face="bold"),
        axis.title.x=element_blank(),
        axis.title.y=element_blank(),
        title=element_text(size=18,face="bold"),
        plot.title = element_text(hjust = 0.5),
        panel.spacing = unit(0, "lines"), 
        strip.background = element_blank(),
        strip.placement = "outside", 
        strip.text =element_text(size=14, face="bold", angle=0))

fig4
```


# 3. combine graphs and text elements

You can put several graphs, textelements and other figures together in an certain order and arrangement using arrangeGrob.

Slightly change the previous graphs
```{r}

fig1 <- ggplot(OrchardSprays, aes(treatment, decrease, colour=factor(treatment)))+
  geom_boxplot()+
  ylim(ymin=0, ymax=150)+
  theme(legend.position='none')

fig2 <- ggplot(OrchardSprays, aes(treatment, rowpos*colpos, colour=factor(treatment)))+
  geom_boxplot()+
  ylim(ymin=0, ymax=150)+
  theme(legend.position='none')

fig3 <- ggplot(OrchardSprays, aes(treatment, colpos, colour=factor(treatment)))+
  geom_boxplot()+
  theme(legend.position='none')


# horizontal boxplots with coord_flip()
labels <- c("1" = "Block 1", "2"= "Block 2", "3" = "Block 3", "4"= "Block 4", "5"= "Block 5", "6"= "Block 6")

fig4 <- ggplot(npk, aes(N, yield, colour=factor(N)))+
  geom_boxplot()+
  coord_flip()+
  labs(title="block effects on yield for N-treatment", y= "yield")+
  facet_wrap(~block,nrow=3,strip.position = "left", scales = "free_y",
             labeller=labeller(block = labels))+
  theme(axis.text=element_text(size=10, face="bold", angle=0),
        axis.title=element_text(size=12,face="bold"),
        axis.title.x=element_blank(),
        axis.title.y=element_blank(),
        title=element_text(size=14,face="bold"),
        plot.title = element_text(hjust = 0.5),
        panel.spacing = unit(0, "lines"), 
        strip.background = element_blank(),
        strip.placement = "outside", 
        strip.text =element_text(size=10, face="bold", angle=0))

```


```{r}
# create a layout as you want it where each number is a separate Grob (figure or textgrob)

lay <- rbind(c(1,NA,NA,2,NA,NA,3,NA,NA),
             c(4,4,4,5,5,5,6,6,6),
             c(4,4,4,5,5,5,6,6,6),
             c(4,4,4,5,5,5,6,6,6),
             c(4,4,4,5,5,5,6,6,6),
             c(4,4,4,5,5,5,6,6,6),
             c(4,4,4,5,5,5,6,6,6),
             c(4,4,4,5,5,5,6,6,6),
             c(4,4,4,5,5,5,6,6,6),
             c(7,NA,NA,NA,NA,NA,NA,NA,NA),
             c(10,10,10,10,10,10,10,10,10),
             c(10,10,10,10,10,10,10,10,10),
             c(10,10,10,10,10,10,10,10,10),
             c(10,10,10,10,10,10,10,10,10),
             c(10,10,10,10,10,10,10,10,10),
             c(10,10,10,10,10,10,10,10,10),
             c(10,10,10,10,10,10,10,10,10),
             c(10,10,10,10,10,10,10,10,10),
             c(10,10,10,10,10,10,10,10,10),
             c(10,10,10,10,10,10,10,10,10))

# define textgrobs
t_a <- textGrob("a")
t_b <- textGrob("b")
t_c <- textGrob("c")
t_d <- textGrob("d")


g_1 <- arrangeGrob(t_a, t_b, t_c, fig1, fig2, fig3, t_d, fig4, layout_matrix = lay) # write the grobs in the correct order
grid.newpage()
grid.draw(g_1)
```

# adding text to a graph with annotate within you chart area
```{r}
ggplot(OrchardSprays, aes(treatment, decrease, colour=factor(treatment)))+
  geom_boxplot()+
  ylim(ymin=0, ymax=150)+
  theme(legend.position='none')+
  annotate("text",x=1,y=150,label="a",size=6)

```

